---
title: "The Effect of Lead Actor's Gender and Movie Title on Gross Earnings"
author: 'Group 4 - Ronald Lee, Lawrence Jiang, Adam Weintraut, Victor Ramirez'
date: "April 4, 2022"
output:
  pdf_document:
    toc: true
    toc_depth: 3
urlcolor: blue    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, echo=FALSE}

library(dplyr)
library(ggplot2)
library(tidyr)
library(scales) 

```

```{r warning=FALSE, message=FALSE, echo=FALSE}

movies_cast_gender <- read.csv("datasets/movies_cast_gender.csv")

movies_cast_gender <- movies_cast_gender %>%
  mutate(gender_str = case_when(
    gender == 1 ~ "female",
    gender == 2 ~ "male"
  ))

movies_dataset <- movies_cast_gender[, c("score", "year","votes","budget","gross", "net_profit",
                                        "runtime", "gender", "genre_", "rating_", "director_",
                                        "writer_", "star_", "country_", "company_",
                                        "female_indicator")]

movies_dataset <- movies_dataset[!is.na(movies_dataset$score), ]
movies_dataset <- movies_dataset[!is.na(movies_dataset$year), ]
movies_dataset <- movies_dataset[!is.na(movies_dataset$votes), ]
movies_dataset <- movies_dataset[!is.na(movies_dataset$budget), ]
movies_dataset <- movies_dataset[!is.na(movies_dataset$gross), ]
movies_dataset <- movies_dataset[!is.na(movies_dataset$runtime), ]
```

# 1. Introduction

## 1.1 Motivation
Movies have been in existence since 1888. Since then, watching movies has been one of our favorite pastimes. Movies are a form of blending visual and sound communication to tell a vivid story. The movie entertainment industry is a high profile multi-billion-dollar global industry.

The ability to predict movie revenue can be a very insightful opportunity. With the predicted revenue information, movie makers can intelligently plan their movie budget. Movie budget line items including star salaries, production, and distribution costs can be smartly negotiated and set. 

The success or failure of a movie depends on a variety of different factors: star cast, cast gender, budget, and title. With the breadth of data available today, making accurate revenue predictions is extremely difficult. However, we do have many data science tools and methodologies at our disposal to help in attempting to make movie revenue predictions. 

## 1.2 Research Question

Our research question is:\
\
_How the Lead actor's gender affect a movie's gross revenue?_
\
\
We have an hypothesis that a male lead actor will bring more gross revenue for a movie overall. In our research project we designed three linear regression models to predict the movie revenue. We joined various movies dataset to extract the following features, cast gender, budget, runtime and title. After data cleaning, data processing and focusing on the action genre we can focus on 1618 movies for analysis. (TODO add analysis success criteria)  

# 2. Description of the Data and Research Design

After you have presented the introduction and the concepts that are under investigation, what data are you going to use to answer the questions? What type of research design are you using? What type of models are you going to estimate, and what goals do you have for these models?  

We use two datasets for our analysis:

**IMDb dataset** ([url](https://www.kaggle.com/datasets/danielgrijalvas/movies)):
For this dataset, the data that we are interested in are the movie title, budget, gross, and runtime. Budget and Gross are in US dollar. Movie Runtime is in minutes.

**Movie roles by gender** ([url](https://www.kaggle.com/code/martinellis/exploration-of-movie-roles-by-gender/notebook)):
This dataset has the gender info for movie actors. The source has instructions on how to execute R code to parse the original data file, which has data in JSON format, and convert it into CSV format that can be imported to R studio. The values of gender is either a value of 2 for "Male", or 1 for "Female".

With this two datasets, we then perform an inner-join on them by their movie lead actor name to generate a single dataset that has all the movie info plus lead actor's gender info.

Before we perform our analysis, we also clean up the data by removing all the entries that have empty values for the fields that we are interested (score, budget, gross, runtime). The final dataset has over 5000 rows of movie data.

## 2a. Data Characteristics
In this section we take a look at the data characteristics for our variables that we are interested and see if there is any need to perform cleanup or transformation before we build our models.\
\

```{r , out.width="70%", warning=FALSE, message=FALSE, echo=FALSE}
ggplot(movies_dataset, aes(budget)) + 
  geom_histogram(position="dodge", alpha=0.8) +
  ggtitle("Distribution of Movie Budgets")+
  scale_x_continuous(labels = comma)
```
\
Movie Budget is highly skewed to the left which means most movies have relatively low budget, and there are only a few movies that have really high budget.
\
```{r , out.width="70%", warning=FALSE, message=FALSE, echo=FALSE}
ggplot(movies_dataset, aes(gross)) + 
  geom_histogram(position="dodge", alpha=0.8) +
  ggtitle("Distribution of Movie Gross Earnings")+
  scale_x_continuous(labels = comma)
```
\
Movie Gross Earnings has similar characteristics as the budget, also highly skewed to the low end.
\
```{r , out.width="70%", warning=FALSE, message=FALSE, echo=FALSE}
ggplot(movies_dataset, aes(runtime)) + 
  geom_histogram(position="dodge", alpha=0.8) +
  ggtitle("Distribution of Movie Runtime")+
  scale_x_continuous(labels = comma)
```
\
Most movies have runtime under 150 minutes.
\

```{r , out.width="70%", warning=FALSE, message=FALSE, echo=FALSE}
ggplot(movies_dataset, aes(gender)) + 
  geom_histogram(position="dodge", alpha=0.8, binwidth = 0.5) +
  ggtitle("Lead Actor's Gender")

```
\
In this histogram that shows the number of Male versus Female Lead Actor. The value "2.0" represent 'Male', and you can see that there are much more Male lead actors in our movie dataset.
\


```{r warning=FALSE, message=FALSE, echo=FALSE}
#movies_cast_gender %>% group_by(gender, gender_str) %>% summarise(count_genders = n()) %>% arrange(desc(count_genders))
# movies_cast_gender %>% group_by(genre, genre_) %>% summarise(count_genres = n()) %>% arrange(desc(count_genres))
# movies_cast_gender %>% group_by(rating, rating_) %>% summarise(count_rating = n()) %>% arrange(desc(count_rating))
# 
# movies_cast_gender %>% group_by(director, director_) %>% summarise(count_directors = n()) %>% arrange(desc(count_directors))
# movies_cast_gender %>% group_by(writer) %>% summarise(count_writers = n()) %>% arrange(desc(count_writers))
# movies_cast_gender %>% group_by(star) %>% summarise(count_stars = n()) %>% arrange(desc(count_stars))
# movies_cast_gender %>% group_by(country) %>% summarise(count_country = n()) %>% arrange(desc(count_country))
# movies_cast_gender %>% group_by(company) %>% summarise(count_company = n()) %>% arrange(desc(count_company))

```




# 3. A Model Building Process

You will next build a set of models to investigate your research question, documenting your decisions. Here are some things to keep in mind during your model building process:

1. *What do you want to measure*? Make sure you identify one, or a few, variables that will allow you to derive conclusions relevant to your research question, and include those variables in all model specifications. How are the variables that you will be modeling distributed? Provide enough context and information about your data for your audience to understand whatever model results you will eventually present. 











2. What [covariates](https://en.wikipedia.org/wiki/Dependent_and_independent_variables#Statistics_synonyms) help you achieve your modeling goals? Are there problematic covariates?  either due to *collinearity*, or because they will absorb some of a causal effect you want to measure?
3. What *transformations*, if any, should you apply to each variable? These transformations might reveal linearities in the data, make our results relevant, or help us meet model assumptions.
4. Are your choices supported by exploratory data analysis (*EDA*)? You will likely start with some general EDA to *detect anomalies* (missing values, top-coded variables, etc.). From then on, your EDA should be interspersed with your model building. Use visual tools to *guide* your decisions. You can also leverage statistical *tests* to help assess whether variables, or groups of variables, are improving model fit.

At the same time, it is important to remember that you are not trying to create one perfect model. You will create several specifications, giving the reader a sense of how robust (or sensitive) your results are to modeling choices, and to show that you're not just cherry-picking the specification that leads to the largest effects.

At a minimum, you need to estimate at least three model specifications: 

The first model you include should include *only the key variables* you want to measure. These variables might be transformed, as determined by your EDA, but the model should include the absolute minimum number of covariates (usually zero or one covariate that is so crucial it would be unreasonable to omit it).

Additional models should each be defensible, and should continue to tell the story of how product features contribute to product success. This might mean including additional right-hand side features to remove omitted variable bias identified by your casual theory; or, instead, it might mean estimating a model that examines a related concept of success, or a model that investigates a heterogeneous effect. These models, and your modeling process should be defensible, incremental, and clearly explained at all points.

Your goal is to choose models that encircle the space of reasonable modeling choices, and to give an overall understanding of how these choices impact results.

# 4. Results

```{r warning=FALSE, message=FALSE, echo=FALSE}

lm = lm(log(gross) ~ female_indicator, movies_dataset)
summary(lm)

lm = lm(log(gross) ~ female_indicator + runtime, movies_dataset)
summary(lm)

lm = lm(log(gross) ~ female_indicator + runtime + budget, movies_dataset)
summary(lm)

```

You should display all of your model specifications in a regression table, using a package like [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) to format your output. It should be easy for the reader to find the coefficients that represent key effects near the top of the regression table, and scan horizontally to see how they change from specification to specification. Make sure that you display the most appropriate standard errors in your table.

In your text, comment on both *statistical significance and practical significance*. You may want to include statistical tests besides the standard t-tests for regression coefficients. Here, it is important that you make clear to your audience the practical significance of any model results. How should the product change as a result of what you have discovered? Are there limits to how much change you are proposing? What are the most important results that you have discovered, and what are the least important? 

# 5. Limitations of your Model 

## 5a. Statistical limitations of your model

As a team, evaluate all of the large sample model assumptions. However, you do not necessarily want to discuss every assumption in your report. Instead, highlight any assumption that might pose significant problems for your analysis. For any violations that you identify, describe the statistical consequences. If you are able to identify any strategies to mitigate the consequences, explain these strategies. 

Note that you may need to change your model specifications in response to violations of the large sample model. 

## 5b. Structural limitations of your model

What are the most important *omitted variables* that you were not able to measure and include in your analysis? For each variable you name, you should *reason about the direction of bias* caused by omitting this variable and whether the omission of this variable calls into question the core results you are reporting. What data could you collect that would resolve any omitted variables bias? 

# 6. Conclusion

Make sure that you end your report with a discussion that distills key insights from your estimates and addresses your research question. 
